---
title: "airbnb_explore_hannah"
author: "Hannah Wen"
date: "2024-05-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NOTE: Run all chunks in order. 

```{r}
library(ggplot2)
library(tidyverse)
library(corrplot)
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
library(ranger)
library(fastDummies)
library(glmnet)
library(corrplot)
library(gt)
library(xgboost)
library(Ckmeans.1d.dp)
library(readr)
#install.packages('osmdata')
#install.packages('sf')
library(osmdata)
library(sf)
library(tigris)
library(leaflet)
library(tidycensus)
```

```{r}
# will need to change pathways for csvs. 
listings = read_csv("/Users/hannahwen/Desktop/STA141C/STA141C_FINAL_PROJECT/Airbnb_Open_Data.csv")
listings
```
# clean typos, remove dollar signs, fill in NA values, 

```{r}
listings %>% select(neighbourhood) %>% distinct()
```
```{r}
listings %>% select(room_type) %>% distinct()
```


```{r}
listings %>% select(neighbourhood_group) %>% table()
```
```{r}
listings %>% filter(neighbourhood_group == 'brookln' | neighbourhood_group == 'manhatan')
```
```{r}
listings
```
```{r}
new_listings = listings %>% select(-c(id, host_id, host_name, host_identity_verified, 'country code', country, license))
new_listings
```

```{r}
# filling na values with a string 
cleaned_listings = new_listings %>% mutate(house_rules = ifelse(is.na(house_rules), "N/A", house_rules))
cleaned_listings = cleaned_listings %>% mutate(NAME = ifelse(is.na(NAME), "N/A", NAME))

cleaned_listings 
```

```{r}
cleaner_listings = cleaned_listings %>% na.omit() # omitting all rows with any na value 
# converting prices into pure numeric values 
cleaner_listings$price = gsub("[$,]", "", cleaner_listings$price) %>% as.numeric()
cleaner_listings$service_fee = gsub("[$,]", "", cleaner_listings$service_fee) %>% as.numeric()
cleaner_listings
```

```{r}
cleaner_listings$neighbourhood_group[13] = "Brooklyn" # fixing typo 
cleaner_listings
```
# Removing false values 




```{r}
# write csv , I commented out all the write csvs in this RMD
#write.csv(cleaner_listings, "/Users/hannahwen/Desktop/STA141C/STA141C_FINAL_PROJECT/cleaned_airbnb_.csv")
```

# ADDITIONAL CLEANING

```{r}
airbnb = read_csv("/Users/hannahwen/Desktop/STA141C/STA141C_FINAL_PROJECT/STA141C_PROJECT/cleaned_airbnb_.csv")
airbnb
```

# Examining house rules variable and creating new variables out of it. 

```{r}
airbnb = airbnb %>% rename(id = "...1") # renaming id column 
airbnb$last_review_year = as.numeric(substr(airbnb$last_review, nchar(airbnb$last_review) - 3, nchar(airbnb$last_review)))
airbnb$no_smoking = grepl("smoking|smoke", airbnb$house_rules) # create columns which identify house rules which mention smoking 

airbnb$house_rules = airbnb$house_rules %>% tolower() 
airbnb$no_rules = airbnb$house_rules=="n/a" # column marking whether there was a house rules section or not 
table(airbnb$no_rules)

```

```{r}
colnames(airbnb)
```

# looking at minimum nights 
```{r}

summary(airbnb$minimum_nights) # inconsistency in data 
airbnb %>% filter(minimum_nights <=0)
# removing values that don't make sense 
airbnb = airbnb %>% filter(minimum_nights > 0)
```

```{r}

summary(airbnb$minimum_nights)
```

# 365 clean 
```{r}
# removing impossible values 
airbnb%>% filter(availability_365<0)
airbnb = airbnb %>% filter(availability_365>=0)
```



# Last review year
```{r}
table(airbnb$last_review_year)
airbnb %>% group_by(last_review_year, neighbourhood_group) %>% summarise(mean_price = mean(price)) %>% 
  spread(key = last_review_year, value = mean_price)
```
```{r}
airbnb = airbnb %>% filter(last_review_year <= 2023)
```

```{r}
airbnb %>% group_by(last_review_year, neighbourhood_group) %>% summarise(mean_price = mean(price)) %>% 
  spread(key = last_review_year, value = mean_price)
```

# Creating number of bedroom variable and examining it

```{r}
airbnb$NAME = tolower(airbnb$NAME)

extract_number_before_string <- function(description) {
  # Regular expression pattern to match the number (integer or decimal) before the specified string
  pattern <- paste0("(?<=\\b)\\d*\\.?\\d+(?=\\s*(bedroom|br|bdrm?|room|rm|bed|bd|-bedroom|-br|-bdrm?|-rm|-bed)\\b)")
  
  # Find matches in the description
  matches <- gregexpr(pattern, description, perl = TRUE)
  
  # Extract the numeric values from matches
  numbers <- regmatches(description, matches)
  
  # Convert numeric values to numeric type to handle both integers and decimals
  numbers <- as.numeric(unlist(numbers))
  
  # Return the first numeric value found
  if (length(numbers) > 0) {
    return(numbers[1])
  } else {
    return(1)  # Return 1 if no numeric value is found
  }
}

# Example usage:
description <- "spacious brooklyn loft - 2.5-bedroom"
string <- "bedroom"
print(extract_number_before_string(description))  # Output: 2.5

```

```{r}
airbnb$bedrooms = as.numeric(sapply(airbnb$NAME, extract_number_before_string))
```




```{r}
cor(airbnb$bedrooms, airbnb$price)
```


```{r}
airbnb %>% group_by(bedrooms) %>% summarise(mean(price))
```
```{r}
table(airbnb$bedrooms)
table(airbnb$bedrooms) %>% prop.table()
```

# finding outliers and fixing possibly wrong classifications 
```{r}
airbnb$bedrooms[airbnb$bedrooms < 1] = 1

airbnb %>% filter(bedrooms >=6)
airbnb[7489, "bedrooms"] = 1

airbnb[29434, "bedrooms"] = 1
airbnb[29436, "bedrooms"] = 1
airbnb[29438, "bedrooms"] = 1
airbnb[29439, "bedrooms"] = 1

airbnb[35623, "bedrooms"] = 1
airbnb[42334, "bedrooms"] = 1
airbnb[51841, "bedrooms"] = 1
airbnb[68244, "bedrooms"] = 1
airbnb[81933, "bedrooms"] = 1
airbnb[81935, "bedrooms"] = 1
airbnb[81937, "bedrooms"] = 1
airbnb[81938, "bedrooms"] = 1
airbnb[81938, "bedrooms"] = 1

```

It's noticed that there are entries named "bedroom 7 bed a" which were interpreted as having 7 bedrooms but appears actually be a shared room. 

```{r}
table(airbnb$bedrooms)
```

Now I will ensure that there aren't more mistakes of assining 6 bedrooms to a listing that is a "bedroom in 6 bed". 
```{r}
airbnb$bedrooms[airbnb$room_type == "Shared"] <- 1
airbnb$bedrooms[airbnb$room_type == "Private room"] <- 1
airbnb$bedrooms = round(airbnb$bedrooms)
```

```{r}
# write.csv(airbnb, "/Users/hannahwen/Desktop/STA141C/STA141C_FINAL_PROJECT/airbnb_new.csv", row.names = FALSE)
airbnb
```



# data analysis portion 
```{r}
airbnb = read_csv("/Users/hannahwen/Desktop/STA141C/STA141C_FINAL_PROJECT/airbnb_new.csv")
airbnb
```
```{r}
airbnb
```
# looking at price 
```{r}
summary(airbnb$price)

```
```{r}
ggplot(airbnb, aes(x = price)) +
  geom_density(fill = "grey", alpha = 0.5) +
  labs(title = "Density Plot of Price",
       x = "Price",
       y = "Density") +
  theme_minimal()
```

# looking at neighbourhood 
```{r}
ggplot(data = airbnb, aes(x = neighbourhood_group, fill=neighbourhood_group)) + geom_bar() + labs() + 
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5) + theme_minimal() +
  labs(title = "Neighborhood Group Distribution", fill = "Neighborhood") + xlab("Neighborhood Group") + ylab("Count") 
```

```{r}
table(airbnb$neighbourhood_group) %>% prop.table()
```

# looking at minimum nights 

```{r}
summary(airbnb$minimum_nights)
```

```{r}
airbnb %>% group_by(room_type) %>% summarise("Median Minimum Nights" = median(minimum_nights))%>% 
  gt() %>% cols_label(room_type = "Room Type")
```

```{r}
airbnb %>% 
        group_by(neighbourhood_group) %>% 
        summarise("Median Minimum Nights" = median(minimum_nights)) %>% 
  gt() %>% cols_label(neighbourhood_group = "Neighborhood Group")
```

```{r}
airbnb %>% group_by(neighbourhood_group) %>% summarise(mean(minimum_nights))
```
# room type 
```{r}
ggplot(data = airbnb, aes(x = room_type, fill=room_type)) + geom_bar() + labs() + geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5) + theme_minimal() +
  labs(title = "Room Type Distribution", fill = "Room Type") + xlab("Room Type") + ylab("Count") 
```

```{r}
table(airbnb$room_type) %>% prop.table()
```
```{r}
table(airbnb$room_type, airbnb$neighbourhood_group) %>% prop.table(margin = 2)
```

# availibility 365
```{r}
summary(airbnb$availability_365)
```

```{r}
airbnb %>% group_by(room_type) %>% summarise("Mean Future Availibility"=mean(availability_365))%>% 
  gt() %>% cols_label(room_type = "Room Type")

```

```{r}
airbnb %>% group_by(neighbourhood_group) %>% summarise("Mean Future Availibility"=mean(availability_365))%>% 
  gt() %>% cols_label(neighbourhood_group = "Neighborhood Group")

```

# construction year 
```{r}
table(airbnb$construction_year)
```

# Examining house rules, cancellation policy, instant_bookable, number of reviews, last review, reviews per month, review rate number, calculated host listing count 

# Cancellation policy Examination 

```{r}
table(airbnb$cancellation_policy)
# frequency table of cancellation policy and instant bookable 
table(airbnb$cancellation_policy, airbnb$instant_bookable)
table(airbnb$cancellation_policy, airbnb$neighbourhood_group) %>% prop.table(margin = 2)
table(airbnb$cancellation_policy, airbnb$no_smoking) %>% prop.table(margin = 2)
table(airbnb$cancellation_policy, airbnb$room_type) %>% prop.table(margin = 2)
```

```{r}
ggplot(data = airbnb, aes(x =neighbourhood_group, fill=cancellation_policy)) + geom_bar(position = "fill") + labs()  + theme_minimal() +
  labs(title = "Cancellation Policies Among Different Neighborhood Groups", fill = "Cancellation Policy Flexibility") + xlab("Neighborhood Group") + ylab("Percent") +
  theme(legend.position = "bottom") +  coord_fixed(ratio = 3.5)
```

# Examining the no smoking variable 
```{r}
table(airbnb$no_smoking) %>% prop.table()
table(airbnb$no_smoking, airbnb$neighbourhood_group) %>% prop.table(margin = 2)
table(airbnb$no_smoking, airbnb$cancellation_policy) %>% prop.table(margin = 2)
airbnb %>% group_by(no_smoking) %>% summarise(mean_price = mean(price))
```
# Looking at the no_rules category 
```{r}
table(airbnb$no_rules) %>% prop.table()
table(airbnb$no_rules, airbnb$neighbourhood_group) %>% prop.table(margin = 2)
table(airbnb$no_rules, airbnb$cancellation_policy) %>% prop.table(margin = 2)
airbnb %>% group_by(no_rules) %>% summarise(mean_price = mean(price))
```


# Mean prices according to cancellation policy and across different neighbourhoodd groups

```{r}

airbnb %>% group_by(cancellation_policy, neighbourhood_group) %>% summarise(mean_price = mean(price)) %>% 
  spread(key = cancellation_policy, value = mean_price) %>% 
  mutate(across(where(is.numeric), round, digits = 2)) %>% gt() %>% 
  cols_label(neighbourhood_group = "Neighbourhood Group", flexible = "Flexible", moderate = "Moderate", 
             strict = "Strict")
airbnb %>% group_by(cancellation_policy) %>% summarise(mean(price))
airbnb %>% group_by(neighbourhood_group) %>% summarise(mean(price))
```

```{r}
ggplot(data = airbnb, aes(x = neighbourhood_group, y = price, color = cancellation_policy)) + geom_boxplot() + theme_minimal() +
  scale_fill_manual(values = c("flexible" = "blue", "moderate" = "green", "strict" = "red")) + 
  labs(x = "Neighbourhood Group", y = "Price", color = "Cancellation Policy", title = "Price Distribution by Cancellation Policy and Neighbourhood Group")

```


```{r}
airbnb %>% group_by(instant_bookable, neighbourhood_group) %>% summarise(mean_price = mean(price)) %>% 
  spread(key = instant_bookable, value = mean_price)
airbnb %>% group_by(instant_bookable) %>% summarise(mean(price))
```

```{r}
ggplot(data = airbnb, aes(x = neighbourhood_group, y = price, color = instant_bookable)) + geom_boxplot() + theme_minimal()+
  scale_fill_manual(values = c("flexible" = "blue", "moderate" = "green", "strict" = "red")) + 
  labs(x = "Neighbourhood Group", y = "Price", color = "Instant Booking Available", title = "Price Distribution by Ability to Book Instantly and Neighbourhood Group")
```

```{r}
# ggplot(data = airbnb, aes(x = reviews_per_month, y = number_of_reviews, color = neighbourhood_group)) + geom_point() + theme_minimal()
cor(airbnb$reviews_per_month, airbnb$number_of_reviews)
cor(airbnb$reviews_per_month, airbnb$review_rate_number)
cor(airbnb$number_of_reviews, airbnb$review_rate_number)
cor(airbnb$reviews_per_month, airbnb$calculated_host_listings_count)
cor(airbnb$number_of_reviews, airbnb$calculated_host_listings_count)
```

```{r}
ggplot(data = airbnb, aes(x = number_of_reviews, y = price), color = neighbourhood) + geom_point() +facet_wrap(~ neighbourhood_group) + theme_minimal()

```


```{r}
colMeans(airbnb %>% dplyr::select(c(reviews_per_month, number_of_reviews, review_rate_number, calculated_host_listings_count)))
```
```{r}
table(airbnb$review_rate_number) %>% prop.table() # distribution of ratings
table(airbnb$review_rate_number, airbnb$neighbourhood_group) %>% prop.table(margin = 2) # distribution of ratings over neighbourhood group
table(airbnb$review_rate_number, airbnb$cancellation_policy) %>% prop.table(margin = 2) # distribution of ratings over cancellation policy 
table(airbnb$review_rate_number, airbnb$no_rules) %>% prop.table(margin = 2) # distribution of ratings over no rules 
median(airbnb$review_rate_number)
```
```{r}
summary(airbnb$calculated_host_listings_count)
```

```{r}

airbnb$review_rate_number_factorized = as.factor(airbnb$review_rate_number)
airbnb %>% group_by(review_rate_number_factorized, neighbourhood_group) %>% summarise(mean_price = mean(price)) %>% 
  spread(key = review_rate_number_factorized, value = mean_price) %>% 
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  gt(caption = "Mean Prices of Each Rating Depending on Neighborhood Group") %>% 
  cols_label(neighbourhood_group = "Neighborhood Group")

```



```{r}
airbnb$review_rate_number_factorized = as.factor(airbnb$review_rate_number)
airbnb %>% group_by(review_rate_number_factorized) %>% summarise(mean_price = mean(price)) %>% 
  spread(key = review_rate_number_factorized, value = mean_price) %>% 
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  gt(caption = "Mean Prices of Each Rating") 
```



```{r}
table(airbnb$bedrooms)
table(airbnb$bedrooms) %>% prop.table()
```
```{r}
airbnb %>% group_by(bedrooms) %>% summarise(mean_price = mean(price)) %>% 
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  spread(key = bedrooms, value = mean_price) %>% gt(caption = "Mean Prices According to Number of Bedrooms") 
```

```{r}
# making number of bedrooms a factor
airbnb$bedrooms = as.factor(airbnb$bedrooms)
```

# examination on construction year
```{r}
airbnb %>% group_by(construction_year) %>% summarise(mean_price = mean(price))
```

```{r}
table(airbnb$construction_year) 
```

```{r}
ggplot(data = airbnb, aes(x =construction_year, fill=as.factor(construction_year))) + geom_bar() + labs()  + theme_minimal() +
  labs(title = "Construction Year Distribution", fill = "Construction Year") + xlab("Construction Year") + ylab("Count") 
```


```{r}
summary(airbnb$construction_year)
```
```{r}
airbnb %>% group_by(room_type) %>% summarise("Median Year"=median(construction_year))%>% 
  gt() %>% cols_label(room_type = "Room Type")
```

```{r}
airbnb %>% group_by(neighbourhood_group) %>% summarise("Median Year"=median(construction_year))%>% 
  gt() %>% cols_label(neighbourhood_group = "Neighborhood Group")
```

```{r}
airbnb_cor = airbnb %>% dplyr::select(-c(id, NAME, neighbourhood, neighbourhood_group, instant_bookable, 
        cancellation_policy, room_type, service_fee, house_rules, no_smoking, no_rules, review_rate_number_factorized,
        last_review))
airbnb_cor$bedrooms = as.numeric(airbnb_cor$bedrooms)
correlation_matrix = cor(airbnb_cor)
corrplot(correlation_matrix, method = "color", tl.cex = 0.8)

```


```{r}
airbnb_cor
```

# clustering

```{r}
cluster_data <- airbnb[, c("lat", "long", "price")]

# Scaling the data - important for clustering
cluster_data_scaled <- scale(cluster_data)
set.seed(123)  # For reproducibility
kmeans_result <- kmeans(cluster_data_scaled, centers = 5, nstart = 25)

# Adding cluster labels to the original data
airbnb$cluster <- as.factor(kmeans_result$cluster)

# Summary statistics for each cluster
library(dplyr)
cluster_summary <- airbnb %>%
  group_by(cluster) %>%
  summarise(
    avg_price = mean(price, na.rm = TRUE),
    median_price = median(price, na.rm = TRUE),
    min_price = min(price, na.rm = TRUE),
    max_price = max(price, na.rm = TRUE),
    count = n(),
    avg_lat = mean(lat),
    avg_long = mean(long)
  )
print(cluster_summary)

ggplot(airbnb, aes(x = long, y = lat, color = factor(cluster))) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd")) +
  labs(title = "Airbnb Clusters Based on Location and Price",
       x = "Longitude",
       y = "Latitude",
       color = "Cluster") +
  theme_minimal()
```
# Map visualizations (WARNING: WILL LAG)

```{r}
# Filter for valid latitude and longitude values
data <- data %>%
  filter(!is.na(lat) & !is.na(long))

# Define colors for different room types
colors <- colorFactor(palette = "viridis", domain = data$room_type)

# Get all county boundaries in New York State
ny_counties <- counties(state = "NY", cb = TRUE)

# Filter for the five boroughs of NYC
nyc_boroughs <- ny_counties %>%
  filter(NAME %in% c("New York", "Kings", "Queens", "Bronx", "Richmond"))

# Ensure the CRS matches
nyc_boroughs <- st_transform(nyc_boroughs, crs = 4326)

# Define colors for different room types
colors <- colorFactor(palette = "viridis", domain = data$room_type)

# Create an interactive map
map <- leaflet(data) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    ~long, ~lat,
    color = ~colors(room_type),
    stroke = FALSE,
    fillOpacity = 0.6,
    radius = 3,
    popup = ~paste(
      "<strong>Room Type:</strong>", room_type, "<br>",
      "<strong>Price:</strong>", price, "<br>",
      "<strong>Minimum Nights:</strong>", minimum_nights, "<br>",
      "<strong>Service Fee:</strong>", service_fee, "<br>",
      "<strong>Availability (365):</strong>", availability_365
    )
  ) %>%
  addPolygons(
    data = nyc_boroughs,
    fillColor = NA, color = "black", weight = 2,
    smoothFactor = 0.5, opacity = 1.0, fillOpacity = 0,
    label = ~NAME
  ) %>%
  addLegend(
    "bottomright", pal = colors, values = ~room_type,
    title = "Room Type",
    opacity = 1
  )

# Print the map
map

```

```{r}
data <- data %>%
  filter(!is.na(lat) & !is.na(long) & !is.na(price))

# Convert Airbnb data to an sf object
airbnb_sf <- st_as_sf(data, coords = c("long", "lat"), crs = 4326)

# Get all county boundaries in New York State
ny_counties <- counties(state = "NY", cb = TRUE)

# Filter for the five boroughs of NYC
nyc_boroughs <- ny_counties %>%
  filter(NAME %in% c("New York", "Kings", "Queens", "Bronx", "Richmond"))

# Ensure the CRS matches
nyc_boroughs <- st_transform(nyc_boroughs, crs = st_crs(airbnb_sf))

# Spatial join to assign boroughs to Airbnb listings
airbnb_with_boroughs <- st_join(airbnb_sf, nyc_boroughs, join = st_intersects)

# Extract the borough name and add it to the dataframe
# we have to add the coordinates back as well
coords <- st_coordinates(airbnb_with_boroughs)

airbnb_with_boroughs <- airbnb_with_boroughs %>%
  mutate(
    borough = NAME.y,
    long = coords[, 1],
    lat = coords[, 2]
  ) %>%
  st_drop_geometry()

# Function to create leaflet map for a given borough
create_borough_map <- function(data, borough_name) {
  # Filter data for the specific borough
  borough_data <- data %>%
    filter(borough == borough_name)
  
  # Define colors for different room types
  colors <- colorFactor(palette = "viridis", domain = borough_data$room_type)
  
  # Create the map
  map <- leaflet(data) %>%
    addProviderTiles(providers$CartoDB.Positron) %>%
    addCircleMarkers(
      ~long, ~lat,
      color = ~colors(room_type),
      stroke = FALSE,
      fillOpacity = 0.6,
      radius = 3,
      popup = ~paste(
        "<strong>Room Type:</strong>", room_type, "<br>",
        "<strong>Price:</strong>", price, "<br>",
        "<strong>Minimum Nights:</strong>", minimum_nights, "<br>",
        "<strong>Service Fee:</strong>", service_fee, "<br>",
        "<strong>Availability (365):</strong>", availability_365
      )
    ) %>%
    addLegend(
      "bottomright", pal = colors, values = ~room_type,
      title = "Room Type",
      opacity = 1
    )
  
  return(map)
}

# List of boroughs
boroughs <- unique(airbnb_with_boroughs$borough %>% na.omit()) 

# Generate and print maps for each borough
for (borough in boroughs) {
  print(borough)
  map <- create_borough_map(airbnb_with_boroughs, borough)
  print(map)
}
```




# MODELING

```{r}
set.seed(4)

model_data = airbnb %>% dplyr::select(-c(service_fee,id, review_rate_number, house_rules, last_review, neighbourhood_group, NAME, cluster))
train_indices <- createDataPartition(model_data$neighbourhood, p = 0.8, list = FALSE)
data_for_train = model_data[train_indices, ]
data_for_test = model_data[-train_indices, ]

```

```{r}
colnames(model_data)
```



# exporting train and test so that the rest of the group can use the same train/test set for modeling

```{r}
data_for_train 
# write.csv(data_for_train, "/Users/hannahwen/Desktop/STA141C/STA141C_FINAL_PROJECT/train_airbnb.csv", row.names = FALSE)
```
```{r}
data_for_test
# write.csv(data_for_test, "/Users/hannahwen/Desktop/STA141C/STA141C_FINAL_PROJECT/test_airbnb.csv", row.names = FALSE)
```
# one hot encoding
```{r}
dummies_model=dummyVars(price~., data = model_data)
model_data_enc = predict(dummies_model, newdata = model_data) %>% as.data.frame()

train_encoded = model_data_enc[train_indices, ]  %>% as.data.frame()
test_encoded = model_data_enc[-train_indices, ] %>%as.data.frame()
train_price = model_data["price"][train_indices, ]
test_price = model_data["price"][-train_indices, ]
```



```{r}
colnames(test_encoded)
```



# Loading Training and Testing Data 

```{r}
train_data = data_for_train
test_data = data_for_test
```


```{r}
var(test_data$price)

```

```{r}
var(test_price$price)
```


# Linear Regression 

```{r}
set.seed(4)
lin_reg = lm( price ~ ., data = train_data)
sum_lin_reg = summary(lin_reg)
```




```{r}
lin_pred = predict(lin_reg, newdata = test_data)
mean( (lin_pred - test_data$price)**2)
var(test_data$price)
```
```{r}
R2(lin_pred, test_data$price)
```

# lasso 
```{r}
set.seed(4)
lambda_values = 10^seq(-2, 2, by = 0.1)
lasso = cv.glmnet(train_encoded %>% as.matrix(), train_price %>% as.matrix(), alpha = 1, lambda = lambda_values)
```

```{r}
lasso_pred <- predict(lasso, newx = test_encoded %>% as.matrix(), s = "lambda.min")
mean( (lasso_pred - test_price$price)**2)
var(test_price$price)
R2(lasso_pred, test_price$price)
```

# factorizing some columns
```{r}
train_data$neighbourhood = as.factor(train_data$neighbourhood)
train_data$bedrooms = as.factor(train_data$bedrooms)
train_data$review_rate_number_factorized = as.factor(train_data$review_rate_number_factorized)
train_data

test_data$neighbourhood = as.factor(test_data$neighbourhood)
test_data$bedrooms = as.factor(test_data$bedrooms)
test_data$review_rate_number_factorized = as.factor(test_data$review_rate_number_factorized)
```


# Random Forest
```{r}
set.seed(4)
# random forest via ranger, using default mtry = p/3 = 17/3
rf_v1 = ranger(price~., data = train_data, importance = "impurity", num.trees = 500, mtry = 5)
summary(rf_v1)
```

```{r}
rf_pred_v1 <- predict(rf_v1, data = test_data)
mean((rf_pred_v1$predictions - test_data$price)**2)

var(test_data$price)
```
```{r}
rf_v1$prediction.error
```


```{r}
mean_true = mean(test_data$price)
R2(rf_pred_v1$predictions, test_data$price)
```


```{r}
par(mar = c(5, 5, 4, 2) + 2)
barplot(rf_v1$variable.importance, main = "Impurity-based Variable Importance",las = 2, cex.names = 0.5, 
        names.arg = names(rf_v1$variable.importance))
```

```{r}
set.seed(4)
# random forest via ranger, using default mtry = p/3 = 17/3
rf_v2 = ranger(price~., data = train_data, mtry = 17, importance = "impurity", num.trees = 500)
summary(rf_v2)
```

```{r}
rf_pred_v2 <- predict(rf_v2, data = test_data)
mean((rf_pred_v2$predictions - test_data$price)**2)
var(test_data$price)
```



```{r}

R2(rf_pred_v2$predictions, test_data$price)
```
```{r}
par(mar = c(5, 5, 4, 2) + 2)
barplot(rf_v2$variable.importance, main = "Impurity-based Variable Importance",las = 2, cex.names = 0.5, 
        names.arg = names(rf_v2$variable.importance))

```

```{r}
set.seed(4)
# random forest via ranger, using default mtry = p/3 = 17/3
rf_v3 = ranger(price~., data = train_data, mtry = 17, importance = "impurity", num.trees = 500, min.node.size = 10)
summary(rf_v3)
```

```{r}
rf_pred_v3 <- predict(rf_v3, data = test_data)
mean((rf_pred_v3$predictions - test_data$price)**2)
var(test_data$price)
R2(rf_pred_v3$predictions, test_data$price)

```



```{r}
set.seed(4)
rf_v4 = ranger(price~., data = train_data, mtry = 17, importance = "impurity", num.trees = 1000)
summary(rf_v4)
```

```{r}
rf_pred_v4 <- predict(rf_v4, data = test_data)
mean((rf_pred_v4$predictions - test_data$price)**2)
var(test_data$price)
R2(rf_pred_v4$predictions, test_data$price)

```

```{r}
# train mse
rf_v4$prediction.error
```

```{r}
par(mar = c(5, 5, 4, 2) + 2)
barplot(rf_v4$variable.importance, main = "Impurity-based Variable Importance",las = 2, cex.names = 0.5, 
        names.arg = names(rf_v4$variable.importance))
```


```{r}
set.seed(4)
# random forest via ranger, using default mtry = p/3 = 16/3
rf_v5 = ranger(price~., data = train_data, mtry = 5, importance = "impurity", num.trees = 1000)
summary(rf_v5)
```

```{r}
rf_pred_v5 <- predict(rf_v5, data = test_data)
mean((rf_pred_v5$predictions - test_data$price)**2)
var(test_data$price)
R2(rf_pred_v5$predictions, test_data$price)

```
```{r}
rf_v5$prediction.error
```


# pca x random forest 



```{r}
set.seed(4)
train_data_pca <- prcomp(train_encoded, 
                         center = TRUE, scale. = TRUE)

# Summarizing PCA results
summary(train_data_pca)
```
```{r}
# Scree plot to visualize variance explained by each principal component
screeplot(train_data_pca, type = "lines", main = "Scree Plot")
```


```{r}
set.seed(4)
pc_scores = as.data.frame(train_data_pca$x[, 1:50])

set.seed(123)  # For reproducibility
pc_test = predict(train_data_pca, newdata = test_encoded)[, 1:50]
price_train =model_data$price[train_indices]
price_test = model_data$price[-train_indices]

rf_pca = ranger(
  formula = price_train ~ .,   # formula for the model
  data = data.frame(pc_scores, price_train),  # combining predictors and response into one data frame
  num.trees = 500,              # number of trees
  mtry = 17,                    # number of variables tried at each split
  verbose = TRUE                # prints the progress and other details
)
```



```{r}

pca_pred = predict(rf_pca, pc_test[, 1:50])
mean((pca_pred$predictions-price_test)**2)
R2(price_test, pca_pred$predictions)
```
```{r}
rf_pca$prediction.error
```



# boosting with xgboost 
```{r}
set.seed(4)
xg1 = xgboost(data = train_encoded %>% as.matrix(), label = train_price %>% as.matrix(), objective = "reg:squarederror", max_depth = 6, nrounds = 50, eta = 0.1)
```
```{r}
xg1_pred = predict(xg1, newdata = as.matrix(test_encoded))
mean((xg1_pred - test_price$price)^2)
R2(xg1_pred, test_price$price)
```


```{r}
set.seed(4)
xg2 = xgboost(data = train_encoded %>% as.matrix(), label = train_price %>% as.matrix(), objective = "reg:squarederror", max_depth = 30, nrounds = 50, eta = 0.1)

```



```{r}

xg2_pred = predict(xg2, newdata = as.matrix(test_encoded))
mean((xg2_pred - test_price$price)^2)
R2(xg2_pred, test_price$price)
```

```{r}
set.seed(4)
xg3 = xgboost(data = train_encoded %>% as.matrix(), label = train_price %>% as.matrix(), objective = "reg:squarederror", max_depth = 50, nrounds = 50, eta = 0.1)

```
```{r}
xg3_pred = predict(xg3, newdata = as.matrix(test_encoded))
mean((xg3_pred - test_price$price)^2)
R2(xg3_pred, test_price$price)
```


```{r}
set.seed(4)
xg4 = xgboost(data = train_encoded %>% as.matrix(), label = train_price %>% as.matrix(), objective = "reg:squarederror", max_depth = 75, nrounds = 50, eta = 0.1)

```
```{r}
xg4_pred = predict(xg4, newdata = as.matrix(test_encoded))
mean((xg4_pred - test_price$price)^2)
R2(xg4_pred, test_price$price)
```

```{r}
set.seed(4)
xg5 = xgboost(data = train_encoded %>% as.matrix(), label = train_price %>% as.matrix(), objective = "reg:squarederror", max_depth = 75, nrounds = 100, eta = 0.01)
```
```{r}
xg5_pred = predict(xg5, newdata = as.matrix(test_encoded))
mean((xg5_pred - test_price$price)^2)
R2(xg5_pred, test_price$price)
```

```{r}
set.seed(4)
xg6 = xgboost(data = train_encoded %>% as.matrix(), label = train_price %>% as.matrix(), objective = "reg:squarederror", max_depth = 100, nrounds = 50, eta = 0.1)
```
```{r}
xg6_pred = predict(xg6, newdata = as.matrix(test_encoded))
mean((xg6_pred - test_price$price)^2)
R2(xg6_pred, test_price$price)
```
```{r}
set.seed(4)
xg7 = xgboost(data = train_encoded %>% as.matrix(), label = train_price %>% as.matrix(), objective = "reg:squarederror", max_depth = 75, nrounds = 100, eta = 0.1)

```

```{r}
set.seed(4)
xg7_pred = predict(xg7, newdata = as.matrix(test_encoded))
mean((xg7_pred - test_price$price)^2)
R2(xg7_pred, test_price$price)
```

```{r}
importance_matrix= xgb.importance(colnames(train_encoded), model = xg4)
xgb.ggplot.importance(importance_matrix[1:15])
```

